---
title: $\chi^2$ Goodness-of-Fit Test
subtitle: for Iris Species
author: Francesco Calzona
format:
  beamer:
    theme: boxes
    colortheme: crane
    font: serif
    header-includes: |
      \definecolor{supergray}{HTML}{f8f8f8}
      \setbeamercolor{background canvas}{bg=supergray}
---

# Math Modeling

## $\chi^2$ distribution
$$X \sim \chi^2_n$$
if and only if
$$X = \sum^n_{i = 1}Z_i^2$$ 
with $Z_i$ independent and
$$Z_i \sim \mathcal{N}(0, 1)$$

## $\chi^2_2$ density function

```{r}
library(ggplot2)
library(latex2exp)

x = c(0:10)
y = dchisq(x, 2)

ggplot(data.frame(x, y), aes(x = x, y=y)) + geom_line(color = "blue") + ylab(TeX("P($\ chi^2_2=x$)"))
```

## $\chi^2_2$ CCDF function

```{r}
library(ggplot2)

x = c(0:10)
y = pchisq(x, 2)

ggplot(data.frame(x, y), aes(x = x, y=1-y)) + geom_line(color="red") + ylab(TeX("P($\ chi^2_2>x$)"))
```

## Approximation to the normal
Let
$$X_i \sim Pois(\mathbf{E}[X_i])$$
If $\mathbf{E}[X_i]$ is big,
$$ \frac{X_i-\mathbf{E}[X_i]}{\sqrt{\mathbf{E}[X_i]}} = \frac{X_i- \mu}{\sigma} \sim \mathcal{N}(0, 1)$$

## Goodness-of-fit test
Assuming that our distribution of assumption $H_0$ is true,  
how likely is it that the population assumes values **equal (or more extreme)** to those of our samples?

## $\chi^2$ Goodness-of-Fit Test {.smaller}
Let $C_i=$ # counts of type $i$,  $n=$# of types.
$$H_0: X_i \sim Pois(E[X_i])$$
for $i \in \{1...n\}$.
The **test statistic** is
$$T_s = \sum_{i=1}^n\frac{(C_i - \mathbf{E}[X_i])^2}{\mathbf{E}[X_i]} \sim \chi^2_{n-1}$$
and
$$P\text{-}value = \mathbf{P}(\chi^2_{n-1} > T_s)$$

## $\chi^2$ Goodness-of-Fit Test for iris dataset
Let $C_i=$ # counts of species $i$, $N=$# of total observations
$$H_0: X_i \sim Pois(\frac{N}{3})$$
$$T_s = \sum_{i=1}^3\frac{(C_i - \frac{N}{3})^2}{\frac{N}{3}}$$
$$P\text{-}value = \mathbf{P}(\chi^2_2 > T_s)$$

# R: Calculating the statistic

## Is our hypothesis plausible?
Let's look into the dataset:
  
```{r}
#| echo: true

data(iris)
summary(iris$Species)
```
## Is our hypothesis plausible?
```{r}
#| echo: false
library(ggplot2)
ggplot(iris, aes(x = Species,  color = Species, fill = Species)) + geom_bar() + theme_classic()
```
## $\chi^2$ Goodness-of-Fit Test in R
**Be careful about the `table` type!**
```{r}
#| echo: true
chisq.test(table(iris$Species), )
```

# We cannot reject the hypothesis of equal frequency.

## Source code
Presentation created with Quarto and Beamer.  



